Smart Drill-Down

### 2015-02-10 TableSummarization/Paper/formal.tex
Changes to paper

We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) subsumes rule ($a$, $b$). Thus, if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$. If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.

\stitle{Rule Lists:}
A {\em rule-list} is an ordered list of rules returned by our system when a smart drill down operation concludes. 
When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

@@

\stitle{Size Weighting Function:} 
$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule. Consider the examples in Table \ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus, the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus, our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).

To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a hundred of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus, this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those tuples. Thus, $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus, this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list. Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.

\stitle{Bits Weighting Function:}
$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$. Like the previous scoring function, this one adds some weight for every non-starred value in a rule. But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the inherent complexity of the column. The reason behind this variation is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules. But this time, we count the number of `bits' that are pre-filled. For a column $c$, specifying a value in the column takes $\lceil \text{log}(|c|)\rceil$ bits of information. A non-starred valued in a rule $r$ thus pre-fills $MCount(r,R) \lceil \text{log}(|c|) \rceil$ bits of the table. Hence this scoring function gives us the the number of bits of the table that are pre-filled by the rule set.

Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.

@@

\begin{table}
\centering
\scriptsize
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\vspace{-10pt}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\vspace{-15pt}
\end{table}

\subsection{Smart drill down Operations}
\label{sec:interface}
When the user starts using a system equipped with
the smart drill down operator, they first see a table with a single trivial rule as shown in table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of table~\ref{table:introexample0} would display table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$. This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.

Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}. This operation is equivalent to a traditional
roll up, but for smart drill downs instead of traditional drill downs.

\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\vspace{-10pt}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\vspace{-10pt}
\end{table}

>>>>

We say that rule $r_1$ is a {\em sub-rule} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) is a sub-rule of ($a$, $b$). If $r_1$ is a sub-rule of $r_2$, then we also say that $r_2$ is a {\em super-rule} of $r_1$. If $r_1$ is a sub-rule of $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$.

\stitle{Rule Lists:}
A {\em rule-list} is an ordered list of rules returned by our system in response to a smart drill down operation. 
When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a new rule-list below $r$.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

@@

\stitle{Size Weighting Function:} 
\techreporttext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$; we define this quantity to be the {\em size} of the rule. Consider the examples in Table \ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus, the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus, our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus, when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).}

\papertext{$W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ : 
Here we set weight equal to the number of non-starred values in the rule $r$ i.e. the {\em size} of the rule. For example, in Table~\ref{table:introexample3}, the rule (Target, bicycles, $\star$) has weight $2$. To get the idea behind using this weighting function, imagine we are trying to reconstruct the table from the rules. We start with a blank value in every column of every table tuple. Then for each displayed rule $r$, we fill each of its values in the corresponding column of $MCount(r)$ tuples. We do this for each rule, filling a total of $\sum_{r} MCount(r)W(r)$ values. Thus maximising score leads to choosing the set of rules that will pre-fill as much of the table as possible.}

\techreporttext{To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a hundred of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus, this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those tuples. Thus, $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus, this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list. Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.}

\stitle{Bits Weighting Function:}
$W(r) = \sum_{c \in C : r(c) \neq \star} \lceil \text{log}_2(|c|) \rceil$ where $|c|$ refers to the number of distinct possible values in column $c$. This scoring function weighs each column depending on its inherent complexity, instead of equally like the Size function. The reason behind is: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Except that this time, we count the number of `bits' of the table that are pre-filled by the rules. This is because for column $c$, specifying a value in the column takes $\lceil \text{log}(|c|)\rceil$ bits of information. This scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.

@@

\techreporttext{
\begin{table}
\centering
\scriptsize
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\vspace{-10pt}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\vspace{-15pt}
\end{table}
}

\subsection{Smart drill down Operations}
\label{sec:interface}
When the user starts using a system equipped with
the smart drill down operator, they first see a table with a single trivial rule as shown in Table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$. This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. \techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.} This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise. Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. \techreporttext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}. This operation is equivalent to a traditional
roll up, but for smart drill downs instead of traditional drill downs.}

\techreporttext{
\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\vspace{-10pt}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\vspace{-10pt}
\end{table}
}

### 2015-02-10 TableSummarization/Paper/intro.tex
Changes to paper

In addition to presenting this new operator called smart drill down, in this paper, 
we also present novel sampling techniques to compute the results for this
operator {\em in an interactive fashion} on increasingly larger databases. 
Unlike the traditional OLAP setting, these computations 
require no pre-materialization, and can be implemented 
within or on top of any relational database system.

>>>>

In addition to presenting this new operator called smart drill down, we also present novel sampling techniques to compute the results for this
operator {\em in an interactive fashion} on increasingly larger databases. 
Unlike the traditional OLAP setting, these computations 
require no pre-materialization, and can be implemented 
within or on top of any relational database system.

### 2015-02-10 TableSummarization/Paper/related.tex
Changes to paper

\section{Related Work}\label{sec:related}
%TODO: Repeat what we said in intro about interactive, etc.
There has been work on finding cubes to browse in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}. This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution. The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones.

@@

We use sampling to find approximate estimates of rule counts. Various other database systems~\cite{Acharya:1999:AAQ:304182.304581, Agarwal:2013:BQB:2465351.2465355} use samples to find approximate results to SQL aggregation queries. These systems create samples in advance and only update them when the database changes. In contrast, we keep updating our samples on the fly, as the user interacts with our system.

% Our algorithm uses ideas from the a priori algorithm~\cite{apriori}.

% Related Work summary : Is ours the only only that is flexible (due to weight function black box)?
% Explanation tables takes lots of papers and talks about them in one go in related work. saying they optimize for different opbjective functions. We can say that as well. We have a unique idea of flexible weighting functions for managing the tradeoff between coverage and specificity. 
% Succintly summarizing with itemsets-- Mampaey:2011:TMI:2020408.2020499 -- Also seemed to optimize according to max entropy criterion
% Efficient and Effective... Tableaux -- Golab_efficientand -- Similar to conditional functional dep. paper.
% Near Optimal Tableaux Conditional functional dep. -- Golab:2008:GNT:1453856.1453900 -- Finding a smallest sized tableau that provides some confidence and support for a conditional functional dependency.
% MDL Summarization with holes -- Bu:2005:MSH:1083592.1083644 -- Min sized summary with a rectangle and exceptions (holes). 
% Generalized MDL Approach -- Lakshmanan:2002:GMA:1287369.1287435 -- Similar to MDL Approach.
% Overlapped Hyperrectangle -- Xiang_succinctsummarization -- Efficiently find the set of hyperrectangles that covers the data (sets of itemsets)
% Tiling Databases Geerts04tilingdatabases Simlar to hyper-rectangles. Probabilistic interpretation of frequent itemsets.

>>>>

\section{Related Work}\label{sec:related}

There has been work on finding cubes to browse in OLAP systems~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration}. This work, along with other existing work~\cite{Mampaey:2011:TMI:2020408.2020499} focuses on finding values that occur more often or less often that expected from a max-entropy distribution. The work does not guarantee good coverage of the table, since it rates infrequently occurring sets of values as highly as frequently occurring ones. Some other data exploration work~\cite{sellam:meet} focuses on finding attribute values that divide the database in equal sized parts, while we focus on values that occur as frequently as possible.

@@

We use sampling to find approximate estimates of rule counts. Various other database systems~\cite{Acharya:1999:AAQ:304182.304581, Agarwal:2013:BQB:2465351.2465355} use samples to find approximate results to SQL aggregation queries. These systems create samples in advance and only update them when the database changes. In contrast, we keep updating our samples on the fly, as the user interacts with our system. There is work on using weighted sampling~\cite{6691587} to create samples favouring data that is of interest to a user, based on the user's history. In contrast, we create samples at run time in response to the user's commands.
